---
title: "R Linear Models"
output: html_notebook
---


```{r}
library(rmarkdown)
this.dir <- paste(rstudioapi::getSourceEditorContext()$path,"/..")
setwd(this.dir)
```

```{r}
inputData = read.csv(file="prostate_cancer.csv", header=TRUE, sep = ",")
attach(inputData)

#inputData[2]:psa~inputData[9]:gleason
logPSA = log(psa)
```
## find the best transformation for target data(log)
## filt out the bad data with p-value
## keep cancervol benpros gleason vesinv
cancervol    0.069454   0.014624   4.749 7.77e-06 ***
benpros      0.087470   0.029605   2.955  0.00401 ** 
gleason      0.358153   0.127976   2.799  0.00629 ** 
vesinv       0.782623   0.268339   2.917  0.00448 ** 
```{r}
#allFit = lm(psa~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
allFit = lm(psa~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
print(summary(allFit))
#allFit1 = lm(logPSA~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
allFit1 = lm(logPSA~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
print(summary(allFit1))
#allFit2 = lm(1/psa~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
#print(summary(allFit2))
#allFit3 = lm(exp(psa)~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
#print(summary(allFit3))
```

Fit3 R-squared is slightly better than Fit2
But, p-value show that we have alot of bad parameter.
And we should keep model simple.
So, Fit2 is the best modle.
```{r}
Fit1 = lm(logPSA~cancervol+benpros+gleason+factor(vesinv))
print(summary(Fit1))
Fit2 = lm(logPSA~log(cancervol)+benpros+gleason+factor(vesinv))
print(summary(Fit2))
Fit3 = lm(logPSA~log(cancervol)+log(cancervol*gleason)+benpros+gleason+factor(vesinv))
print(summary(Fit3))
```
## let's estimate the mean psa
   (Intercept) log(cancervol)        benpros        gleason         vesinv 
   -0.31632980     0.50503308     0.06418978     0.26293703     0.65880768 
```{r}
Fit2$coefficients
est.psa.log = -0.31632980 + 0.50503308*mean(log(cancervol)) + 0.06418978*mean(benpros)+0.26293703*mean(gleason)+ 0.65880768*mean(vesinv)
est.psa = exp(est.psa.log)
print(paste("psa.mean:",mean(psa),("est.psa:"),est.psa))
```

# let's check the qq Plot
```{r}
par(mfrow = c(2, 3))
qqnorm(psa)
qqline(psa)
qqnorm(cancervol)
qqline(cancervol)
qqnorm(benpros)
qqline(benpros)
qqnorm(vesinv)
qqline(vesinv)
qqnorm(gleason)
qqline(gleason)

```
## cancervol is the only one suitable for log transformation
We should turn the input data to normal to get the better model.
```{r}
par(mfrow = c(1, 2))
qqnorm(log(cancervol))
qqline(log(cancervol))
qqnorm(1/(cancervol))
qqline(1/(cancervol))
```
R also provide auto parameter selection by AIC value.
The result just like us.
```{r}
fit <- step(lm(logPSA ~ 1), scope = list(lower = ~1, upper = ~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen), direction = "both")
print(summary(fit))

```
We can also use lm.ridge to add penalty term(supress weight)
The weight of age, weight, capspen is small, so we should remove it.
```{r}
library(MASS)
rfit = lm.ridge(logPSA~cancervol+benpros+gleason+factor(vesinv)+age+weight+capspen)
print(rfit$coef)
#weight of age weight capspen extremely samll

```

